---
title: Outro
fontsize: 9pt
output:
  beamer_presentation:
    theme: "Dresden"
    colortheme: "rose"
    fonttheme: "professionalfonts"
---

```{r, echo = F}

set.seed(1)

```

# You, Me, and NSE

Non-standard evaluation is bananas!

# `ggplot2` - Graphics at the spead of thought

Really fast plotting with ggplot2.

```{r}

library(ggplot2)
dat <- data.frame(let = sample(letters[c(1:5,1:3,1:2)],
                               100,
                               replace = T))

```

# `ggplot2` - Graphics at the spead of thought

```{r}

ggplot(dat) +
    geom_bar(aes(x=let, fill = let), stat="count")

```

# tidverse

Write data processing using verb like syntax, chaining together a pipeline of data processing routines.

```{r}

library(tidyverse)
tib <- tibble(x = runif(10000, 1, 100),
              y = runif(10000, 1, 100),
              let=sample(letters[c(1:5,1:3)], 10000, replace = T))
tib %>%
    filter(let %in% c("a", "b", "c")) %>%
    mutate(prod = x*y) -> tib2

```

# data.table

Super fast tabular data manipulation with a very succinct syntax.

```{r}

library(data.table)
dt <- data.table(x = runif(10000, 1, 100),
                 y = runif(10000, 1, 100),
                 let=sample(letters[c(1:5,1:3)], 10000, replace = T))

```

# data.table

Group by using the `by` argument.

```{r}

dt[, .N, by = let]

```

# data.table

Assignment by reference.

```{r}

dt[, prod:=x*y]
dt

```
