---
title: Outro
fontsize: 9pt
output:
  beamer_presentation:
    theme: "Dresden"
    colortheme: "rose"
    fonttheme: "professionalfonts"
---

```{r, echo = F}

set.seed(1)

```

# You, Me, and NSE

Non-standard evaluation is bananas!

```{r}

nseFun <- function(fruit){
    fruit <- deparse(substitute(fruit))
    if(grepl("bananas", fruit, ignore.case = T)){
        return("Bananas!")
    }
    return("Not bananas...")
}

```

```{r}

nseFun(oranges)
nseFun(bananas)

```

# `ggplot2` - Graphics at the spead of thought

Really fast plotting with ggplot2.

```{r}

library(ggplot2)
dat <- data.frame(let = sample(letters[c(1:5,1:3,1:2)],
                               100, replace = T),
                  frm = sample(c("This", "That", "Other"),
                               1000, replace = T))

```

# `ggplot2` - Graphics at the spead of thought

```{r}

ggplot(dat) +
    geom_bar(aes(x = let, fill = let), stat="count") +
    facet_wrap(~frm) +
    theme_bw()

```

# tidyverse

A whole suite of data procesing packages that allow for very readable, easy to write code.

Includes ggplot2. Very easy and intuative for beginners.

Write data processing using verb like syntax, chaining together a pipeline of data processing routines.

```{r}

library(tidyverse)
tib <- tibble(x = runif(10000, 1, 100),
              y = runif(10000, 1, 100),
              let = sample(letters[c(1:5,1:3)],
                           10000, replace = T))

```

# tidyverse

See how nicely it prints?

```{r}

tib

```

# tidyverse

Pipe-lining with magrittr: `%>%`

```{r}

tib %>%
    filter(let %in% c("a", "b", "c")) %>%
    mutate(prod = x*y) -> tib

tib

```

# data.table

Super fast tabular data manipulation with a very succinct syntax.

Building a `data.table` object is just like `data.frame`.

```{r}

library(data.table)
dt <- data.table(x = runif(10000, 1, 100),
                 y = runif(10000, 1, 100),
                 let = sample(letters[c(1:5,1:3)],
                              10000, replace = T))

```

# data.table

Interfacing with the object leverages NSE for quick, succinct coding. 

Group by using the `by` argument, counting the grouped elements.

```{r}

dt[order(let), .N, by = let]

```

# data.table

Assignment by reference.

```{r}

dt[, prod:=x*y]
dt

```
